---
title: "Practice Exam 1"
author: "Phil Anderson"
date: "October 3, 2017"
output: 
  html_document:
  toc: true
  theme: "journal"
#  highlight: tango
---

```{r, message=F}
# Install our third-party modules
library("tidyverse")
```

First, let's read in our data

```{r}
pref <- read.csv("C:/Users/Philip/Schools/TAMU/STAT_636/exam_prep/consumer_pref.csv")

names(pref) <- tolower(names(pref))
str(pref)
```

## Question 1-A
Generate QQ Plots for each variable

```{r}
par(mfrow=c(2,3))
qqnorm(pref$taste); qqline(pref$taste)
qqnorm(pref$value); qqline(pref$value)
qqnorm(pref$flavor); qqline(pref$flavor)
qqnorm(pref$snack); qqline(pref$snack)
qqnorm(pref$energy); qqline(pref$energy)
```

## Question 1-B
Construct a 95% confidence region for the population mean vector mu.  What are its primary axes and associated half-lengths?  Note that the primary axes will be given by the eigenvectors of the matrix's variance-covariance matrix; in this case they are not readily interpretable, given the number of dimensions.  The half lengths will be given by a constant times the eigenvalues of that covariance matrix.

```{r}
(x_bar <- colMeans(pref))
(S <- var(pref))
(p <- ncol(pref))
(n <- nrow(pref))

pref_eig <- eigen(S)
# axes given by:
pref_eig$vectors

#half-lengths given by:
sqrt(pref_eig$values) * sqrt( (p * (n - p)) / (n * (n - p)) * qf(1-0.05,p, n-p )  )
  
```


## Question 1-C
Report 95% Bonferroni-based simultaneous confidence intervals for our p-variables.  How would you interpret this collection of confidence intervals, in terms of the coverage probabilities?

```{r}
# note that we are using alpha=0.025 for the two-sided coverage
# also note that we are taking the diagonal elements of the variance-covariance matrix

bon_mat <- matrix(NA, nrow=p, ncol=2)

for (i in 1:p) {
bon <- x_bar[i] + c(-1, 1) * qt(1 - 0.025 / p, n - 1) * sqrt(S[i,i] / n)
bon_mat[i, ] <- bon
  }

cbind(bon_mat[,1], x_bar, bon_mat[,2])
```

These intervals are constructed such that we are 95% confident that they contain the true means.  Note that this is not saying that we have a 95% probability that the true mean is contained within the interval.

## Question 1-D
Carry out Hotelling's multivariate generalization of the t-test for the null hypothesis that the mean vector is equal to the null vector.

```{r}
(mu_0 <- rep(0,p))

# gather our test statistic (observed)
(T2_obs <- n * t(x_bar - mu_0) %*% solve(S) %*% (x_bar - mu_0))

# gather our p-value
1 - pf( (n - p) / ((n - 1) * p) * T2_obs , p, n-p )

```

We do not see evidence that our p-value is significantly different from the zero vector.


## Question 1-E
Use the bootstrap algorithm to compute the p-value from the previous problem.

```{r}
B <- 1000 # number of epochs
set.seed(101)

# recall that T2_obs is our observed test statistic for this procuedure

# rescale our data vector
pref_0 <- scale(pref, center=T, scale=F)
T2_output <- numeric(B)

for (i in 1:B){
  temp <- pref_0[sample(1:n, replace=T),]
  x_bar_temp <- colMeans(temp)
  S_temp <- var(temp)
  T2_b <- drop(n * t(x_bar_temp - mu_0) %*% solve(S_temp) %*% (x_bar_temp - mu_0))
  T2_output[i] <- T2_b
}

# let's view our sampling distribution
plot(density(T2_output), main="T2 Null Sampling Distribution")
mtext("Observed Test Statistic in Red")
abline(v=drop(T2_obs), col="red")
sum(T2_output >= drop(T2_obs)) / 1000

```

